---
title: "Assignment 1a (Python) - refactor obesity dataset code"
output: html_notebook
---

The below code downloads data from the obesity dataset from the [UC Irvine ML repository](https://doi.org/10.24432/C5H31Z). It then preprocesses the data, fits a machine-learning model and shows results. The coding style intentionally is not perfect.

In this assignment you'll refactor this code using the principles taught in the Reproducible Coding practices course. Things to consider are for example:

* Modularity
* Documentation
* Tests
* Packaging
* Dependencies
* Version control
* etc.

For the full assessment criteria see the Rubric. For full details on the assignment see the `README.md`

```{r}
install.packages('httr', repos='http://cran.us.r-project.org')
install.packages('readr', repos='http://cran.us.r-project.org')
install.packages('ggplot2', repos='http://cran.us.r-project.org')
install.packages('randomForest', repos='http://cran.us.r-project.org')

library(httr)
library(readr)
library(ggplot2)
library(randomForest)
```

```{r}
url <- 'https://archive.ics.uci.edu/static/public/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition.zip'

r <- GET(url)

tmp <- GET(url, write_disk('data.zip', overwrite = TRUE))
stop_for_status(tmp)

unzip('data.zip', exdir = './data')

df <- read_csv('data/ObesityDataSet_raw_and_data_sinthetic.csv', show_col_types = FALSE)
```

```{r}
df_web <- df[1:498, ]
df_synthetic <- df[499:nrow(df), ]
df_web
```

```{r}
X_columns <- c('Gender', 'Age', 'BMI')

df_web$BMI <- df_web$Weight / (df_web$Height)^2
df_synthetic$BMI <- df_synthetic$Weight / (df_synthetic$Height)^2

plot(df_web$BMI, type = 'l')

outcomes <- c()
for (bmi in df_web$BMI) {
  if (bmi < 18.5) {
    outcomes <- c(outcomes, '0_undwerweight')
  } else if (bmi < 25) {
    outcomes <- c(outcomes, '1_normal')
  } else if (bmi < 30) {
    outcomes <- c(outcomes, '2_overweight')
  } else {
    outcomes <- c(outcomes, '3_obese')
  }
}

df_web$label <- outcomes
```

```{r}
df_web[10:20, ]
```

```{r}
print(colnames(df))

df_web$MTRANS[df_web$MTRANS == 'Walking'] <- 1
df_web$MTRANS[df_web$MTRANS == 'Bike'] <- 1
df_web$MTRANS[df_web$MTRANS == 'Public_Transportation'] <- 0
df_web$MTRANS[df_web$MTRANS == 'Automobile'] <- 0
df_web$MTRANS[df_web$MTRANS == 'Motorbike'] <- 0
df_web$MTRANS <- as.numeric(df_web$MTRANS)

df_web$Gender[df_web$Gender == 'Male'] <- 1
df_web$Gender[df_web$Gender == 'Female'] <- 0
df_web$Gender <- as.numeric(df_web$Gender)

Xcol_binary <- c(
  'family_history_with_overweight',
  'FAVC',
  'SMOKE',
  'SCC'
)

for (col in Xcol_binary) {
  df_web[df_web[[col]] == 'yes', col] <- '1'
  df_web[df_web[[col]] == 'no', col] <- '0'
  df_web[[col]] <- as.numeric(df_web[[col]])
}

Xcol_ordinal <- c('CAEC', 'CALC')

df_web$CAEC <- as.numeric(factor(df_web$CAEC, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1
df_web$CALC <- as.numeric(factor(df_web$CALC, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1
```

```{r}
X_columns <- c('Age', 'Gender', 'family_history_with_overweight',
               'FAVC', 'FCVC', 'NCP', 'CAEC', 'SMOKE', 'CH2O', 'SCC', 'FAF', 'TUE',
               'CALC', 'MTRANS')
```

## Regression problem

```{r}
classes <- sort(unique(df_web$label))
freq <- table(df_web$label)
class_weight <- length(df_web$label) / (length(classes) * as.numeric(freq))
names(class_weight) <- names(freq)

print(classes)
print(class_weight)
```

```{r}
# Note: result can deviate from the python code due to random seed!
set.seed(42)
idx <- seq_len(nrow(df_web))

idx_train <- sample(idx, size = floor(0.7 * length(idx)))
idx_test <- setdiff(idx, idx_train)

X_train <- df_web[idx_train, X_columns]
X_test <- df_web[idx_test, X_columns]
y_train <- df_web$BMI[idx_train]
y_test <- df_web$BMI[idx_test]

sample_weights <- class_weight[df_web$label[idx_train]]
```

```{r}
reg <- lm(BMI ~ ., data = cbind(X_train, BMI = y_train))
reg <- randomForest(x = X_train, y = y_train, ntree = 100, nodesize = 5, maxnodes = 20)
```

```{r}
preds <- predict(reg, X_test)

plot(preds, y_test, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
plot(predict(reg, X_train), y_train, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
sample_weights_test <- class_weight[df_web$label[idx_test]]

mae_baseline <- sum(abs(y_test - rep(mean(y_train), length(y_test))) * sample_weights_test) / sum(sample_weights_test)
mae_model <- mean(abs(y_test - preds) * sample_weights_test)

cat(sprintf('Mean absolute error (baseline): %.3f\n', mae_baseline))
cat(sprintf('Mean absolute error: %.3f\n', mae_model))
```
