---
title: "Assignment 1a (Python) - refactor obesity dataset code"
output: html_notebook
---

The below code downloads data from the obesity dataset from the [UC Irvine ML repository](https://doi.org/10.24432/C5H31Z). It then preprocesses the data, fits a machine-learning model and shows results. The coding style intentionally is not perfect.

In this assignment you'll refactor this code using the principles taught in the Reproducible Coding practices course. Things to consider are for example:

* Modularity
* Documentation
* Tests
* Packaging
* Dependencies
* Version control
* etc.

For the full assessment criteria see the Rubric. For full details on the assignment see the `README.md`

```{r}
install.packages('httr', repos='http://cran.us.r-project.org')
install.packages('readr', repos='http://cran.us.r-project.org')
install.packages('ggplot2', repos='http://cran.us.r-project.org')
install.packages('randomForest', repos='http://cran.us.r-project.org')

library(httr)
library(readr)
library(ggplot2)
library(randomForest)
```

```{r}
data_url <- 'https://archive.ics.uci.edu/static/public/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition.zip'

#removed this line as it is not needed

tmp <- httr::GET(data_url, write_disk('data.zip', overwrite = TRUE))
httr::stop_for_status(tmp)

unzip('data.zip', exdir = './data')

obesity_data <- readr::read_csv('data/ObesityDataSet_raw_and_data_sinthetic.csv', show_col_types = FALSE)
```

```{r}
survey_data <- obesity_data[1:498, ]
synthetic_data <- df[499:nrow(df), ]
survey_data
```

```{r}
##remove redundancy
##X_columns <- c('Gender', 'Age', 'BMI')

##Create BMI variable based on standard formula
survey_data$BMI <- survey_data$Weight / (survey_data$Height)^2
synthetic_data$BMI <- synthetic_data$Weight / (synthetic_data$Height)^2

#scatterplot of BMI values
ggplot2::plot(survey_data$BMI, type = 'p')

#label data based on standard BMI thresholds
outcomes <- c()
for (bmi in survey_data$BMI) {
  if (bmi < 18.5) {
    outcomes <- c(outcomes, '0_underweight')
  } else if (bmi < 25) {
    outcomes <- c(outcomes, '1_normal')
  } else if (bmi < 30) {
    outcomes <- c(outcomes, '2_overweight')
  } else {
    outcomes <- c(outcomes, '3_obese')
  }
}

survey_data$label <- outcomes
```

```{r}
survey_data[10:20, ]
```

```{r}
##print(colnames(obesity_data))
##remove redundant line

##differentiate active and passive transport with 0/1 for modelling
survey_data$MTRANS[survey_data$MTRANS == 'Walking'] <- 1
survey_data$MTRANS[survey_data$MTRANS == 'Bike'] <- 1
survey_data$MTRANS[survey_data$MTRANS == 'Public_Transportation'] <- 0
survey_data$MTRANS[survey_data$MTRANS == 'Automobile'] <- 0
survey_data$MTRANS[survey_data$MTRANS == 'Motorbike'] <- 0
survey_data$MTRANS <- as.numeric(survey_data$MTRANS)

##convert gender to numeric values for modelling
survey_data$Gender[survey_data$Gender == 'Male'] <- 1
survey_data$Gender[survey_data$Gender == 'Female'] <- 0
survey_data$Gender <- as.numeric(survey_data$Gender)

##convert binary columns to numeric values for modelling
binary_columns <- c(
  'family_history_with_overweight',
  'FAVC',
  'SMOKE',
  'SCC'
)

for (col in binary_columns) {
  survey_data[survey_data[[col]] == 'yes', col] <- '1'
  survey_data[survey_data[[col]] == 'no', col] <- '0'
  survey_data[[col]] <- as.numeric(survey_data[[col]])
}

##Convert ordinal columns to numeric values for modelling
ordinal_columns <- c('CAEC', 'CALC')

for (col in ordinal_columns) {
  survey_data[[col]] <- as.numeric(factor(survey_data[[col]], levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1
}
```

```{r}
X_columns <- c('Age', 'Gender', 'family_history_with_overweight',
               'FAVC', 'FCVC', 'NCP', 'CAEC', 'SMOKE', 'CH2O', 'SCC', 'FAF', 'TUE',
               'CALC', 'MTRANS')
```

## Regression problem

```{r}
weight_classes <- sort(unique(survey_data$label))
freq <- table(survey_data$label)

#calculate class weights to handle class imbalance
class_weight <- length(survey_data$label) / (length(weight_classes) * as.numeric(freq))
names(class_weight) <- names(freq)

print(weight_classes)
print(class_weight)
```

```{r}
# Note: result can deviate from the python code due to random seed!
set.seed(42)
idx <- seq_len(nrow(survey_data))

#choose rows to train and test with
idx_train <- sample(idx, size = floor(0.7 * length(idx)))
idx_test <- setdiff(idx, idx_train)

X_train <- survey_data[idx_train, X_columns]
X_test <- survey_data[idx_test, X_columns]
y_train <- survey_data$BMI[idx_train]
y_test <- survey_data$BMI[idx_test]

sample_weights <- class_weight[survey_data$label[idx_train]]
```

```{r}
#run linear regression because we expect a linear relationships between features and BMI
reg <- lm(BMI ~ ., data = data.frame(X_train, BMI = y_train))
```

```{r}
#evaluate model accuracy in testing set
preds <- predict(reg, X_test)

plot(preds, y_test, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
#evaluate model accuracy in training set
plot(predict(reg, X_train), y_train, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
sample_weights_test <- class_weight[survey_data$label[idx_test]]

#compute baseline mean absolute error
mae_baseline <- sum(abs(y_test - rep(mean(y_train), length(y_test))) * sample_weights_test) / sum(sample_weights_test)

#compute mean absolute error of model
mae_model <- mean(abs(y_test - preds) * sample_weights_test)

cat(sprintf('Mean absolute error (baseline): %.3f\n', mae_baseline))
cat(sprintf('Mean absolute error: %.3f\n', mae_model))
```